Supplemental information

### 2018-01-19 manuscript/supplemental.tex
added some comments and made pretty

\section*{Appendix S1: Obtaining prior interaction probabilities}

@@

      Barbour \emph{et al.} sampled 145 branches from 
      \emph{Salix hookeriana} of 26 genotypes. They recorded galls 
      made by four species of Cecidomyid midges on each branch. We 
      transformed these data to a binary genotype-galler network 
      matrix, where an entry $ij$ was 1 if galler $i$ was observed 
      on any branch of genotype $j$ and 0 otherwise. In total, this 
      network contained 75 realised galler-genotype interactions 
      out of a potential 104. One \emph{S. hookeriana} genotype did
      not interact with any gallers and was removed from the network.
      The R code used to extract this 
      network follows:

      \vspace{12pt}
      \emph{
      \noindent \hspace{-4pt} prior\_web\_data=read.csv(``tree\_level\_interaxn\_all\_plants\_traits\_size.csv")\\
      \# Remove extra data on parasitoids, etc.\\
      prior\_web\_data[,31:57]\textless-NULL\\
      prior\_web\_data[,4:26]\textless-NULL\\
      prior\_web\_data[,2]\textless-NULL\\
      \# Build a web with interactions=1 for any galler observed on any genotype\\
      prior\_web=matrix(nrow=26,ncol=4)\\
      for(r in 1:length(levels(prior\_web\_data\$Genotype)))\{\\
      \hspace{4pt}  gen=levels(prior\_web\_data\$Genotype)[r]\\
      \hspace{4pt}  subset=prior\_web\_data[which(prior\_web\_data\$Genotype==gen),]\\
      \hspace{4pt}  for(c in 1:4)\{\\
      \hspace{8pt}    if(sum(subset[,2+c])\textgreater0)\{\\
      \hspace{12pt}      prior\_web[r,c]=1\\
      \hspace{8pt}    \} else \{\\
      \hspace{12pt}      prior\_web[r,c]=0\\
      \hspace{8pt}    \}\}\}
      \# Remove non-interactive genotype
      prior\_web<-prior\_web[which(rowSums(prior\_web)\textgreater0),]
      }

@@

@@

      \vspace{12pt}
      \emph{
      \noindent \hspace{-4pt} 
      prior\_web\_data=read.csv('tree\_level\_interaxn\_all\_plants\_traits\_size.csv')\\
      prior\_web\_data <- prior\_web\_data[,8:26]

      \# Build a web with interactions=1 for any parasitoid on any galler
      prior\_web=data.frame()

      \# each column contains all interactions for a particular parasitoid-galler combination
      for(column in 1:ncol(prior\_web\_data)){
      \hspace{-4pt}   \# the column name gives the species which makes the gall, followed by the species emerging from the gall
      \hspace{-4pt}   galler\_emerged=strsplit(colnames(prior\_web\_data)[column], "\_")
      \hspace{-4pt}   galler = galler\_emerged[[1]][1]
      \hspace{-4pt}   emerged = galler\_emerged[[1]][2]

      \hspace{-4pt}   \# the strength of the interaction is the sum of the column (number of galls across all branches sampled)
      \hspace{-4pt}   linkstrength = sum(prior\_web_data[,column])
      \hspace{-4pt}   prior\_web[as.character(galler),as.character(emerged)] <- linkstrength
      }

      \# Make the naS into 0
      prior\_web[is.na(prior\_web)] <- 0

      \# remove emerging gallers with no parasitoids
      prior\_web\_para\_only <- prior\_web[,c(2:6,8:9,12)]

      \# convert interaction strengths to 1
      prior\_web[which(prior\_web>0)] <- 1

      write.csv(prior\_web\_para\_only, file="prior\_web\_para\_only.csv", quote=F)
      }

@@

      \vspace{12pt}
      \emph{
      \noindent \hspace{-4pt} \# Now get the degree distributions from the web\\
      deg\_dist\_Salix=rowSums(prior\_web)/ncol(prior\_web)\\
      \# Remove one genotype that never interacted\\
      deg\_dist\_Salix\textless-deg\_dist\_Salix[which(rowSums(prior\_web)\textgreater0)]\\
      deg\_dist\_galler=colSums(prior\_web)/nrow(prior\_web)\\
      \# Interaction probabilities are the product of plant and galler probabilities\\
      int\_probs=as.numeric(deg\_dist\_galler\%$*$\%t(deg\_dist\_Salix))\\

      \# And for the galler-parasitoid web:\\
      deg\_dist\_parasitoid=rowSums(prior\_web\_para\_only)/ncol(prior\_web\_para\_only)\\
      deg\_dist\_parasitized=colSums(prior\_web\_para\_only)/nrow(prior\_web\_para\_only)\\
      gp\_int\_probs=as.numeric(deg\_dist\_parasitized\%$*$\%t(deg\_dist\_parasitoid))

      }

@@

\section*{Appendix S2: Calculating posterior distributions and confidence intervals}

  The following simple functions are implemented in the R language. 
  Priordata is the list of interaction frequencies from the prior 
  data, n is the number of sites with observed co-occurrance of 
  species $i$ and $j$, and k is the number of sites with an 
  observed interaction $ij$. When calculating the prior 
  distribution, both n and k are 0. In the main text we assume 100\%
  confidence in observed interactions and therefore consider only 
  cases where k=0. Only n varies between species pairs.

  \vspace{12pt}
  \emph{
  \noindent \hspace{-4pt}calculate\_parameters\textless-function(priordata,n,k)\{\\
  \hspace{4pt} \# Calculate prior parameters\\
  \hspace{4pt} start=list(shape1=1,shape2=1)
  \hspace{80pt} pars=fitdistr(x=priordata,`beta',start=start,lower=c(0,0))\$estimate\\
  \hspace{4pt} alpha=pars[[1]]\\
  \hspace{4pt} beta=pars[[2]]\\
  \hspace{4pt} \# Update the parameters with data. If n=0 and k=0, no change.\\
  \hspace{4pt} alpha\_prime=alpha+k\\
  \hspace{4pt} beta\_prime=beta+n-k\\
  \hspace{4pt} pars2=c(alpha\_prime,beta\_prime)\\
  \hspace{4pt} return(pars2) \}
  }
  \vspace{12pt}

    To calculate the maximum likelihood estimates of the mean and 
    variance of the probability of interaction $\lambda_{ij}$:

    \vspace{12pt}
    \emph{
    calculate\_distribution\textless-function(pars)\{\\
      \# alpha and beta may be from a prior or posterior distribution\\
      alpha=pars[[1]]\\
      beta=pars[[2]]\\

      \# Calculate the MLE of the mean\\
      mu\_num=alpha\\
      mu\_den=alpha+beta\\
      mu=mu\_num\/mu\_den\\

      \# Calculate the MLE of the variance\\
      sig\_num=alpha\*beta\\
      den1=alpha+beta\\
      den2=den1\*\*2\\
      sig\_den=den1\*den2\\
      sigma2=sig\_num\/sig\_den\\

      return(c(mu,sigma2))\}
    }
  \vspace{12pt}

  \vspace{12pt}
  \emph{
  credible\_interval\textless-function(pars,p\_lower,p\_upper)\{\\
    alpha=pars[[1]]\\
    beta=pars[[2]]\\
    lowCI=qbeta(p=p\_lower,shape1=alpha,shape2=beta)\\
    highCI=qbeta(p=p\_upper,shape1=alpha,shape2=beta)\\
    return(c(lowCI,highCI))\}
  }

  \vspace{12pt}
  \emph{
  plot\_precision\textless-function(threshold,confidence,pars)\{\\
    alpha=pars[[1]]\\
    beta=pars[[2]]\\
    n=seq(0,100,1)\\
    k=0\\
    cdf=pbeta(threshold,shape1=alpha,shape2=beta+n)\\
    samples=length(which(cdf<confidence))\\
    return(samples) \}
  }

    \bibliographystyle{ecollett} 
    \bibliography{manual_abbrev} % Abbreviate journal titles.

>>>>

\section*{Appendix S1: Calculating posterior distributions and confidence intervals}

  The following simple functions are implemented in the R language. They are 
  also provided as R code in a separate file.
  Priordata is the list of interaction frequencies from the prior 
  data, n is the number of sites with observed co-occurrance of 
  species $i$ and $j$, and k is the number of sites with an 
  observed interaction $ij$. When calculating the prior 
  distribution, both n and k are 0. In the main text we assume 100\%
  confidence in observed interactions and therefore consider only 
  cases where k=0. Only n varies between species pairs.

  \vspace{12pt}
  \begin{em}
  \noindent \hspace{2pt}calculate\_parameters\textless-function(priordata,n,k)\{\\
  \vspace{4pt}
    \# Calculate prior parameters\\
    start=list(shape1=1,shape2=1)\\
    pars=fitdistr(x=priordata,`beta',start=start,lower=c(0,0))\$estimate\\
    alpha=pars[[1]]\\
    beta=pars[[2]]\\
  \vspace{4pt}
    \# Update the parameters with data. If n=0 and k=0, no change.\\
    alpha\_prime=alpha+k\\
    beta\_prime=beta+n-k\\
    pars2=c(alpha\_prime,beta\_prime)\\
    return(pars2) \}
  \end{em}
  \vspace{12pt}

    All of the following functions use the parameters returned by 
    ``calculate\_parameters". To calculate the maximum likelihood estimates of 
    the mean and variance of the probability of interaction $\lambda_{ij}$:

    \vspace{12pt}
    \begin{em}
    \noindent \hspace{2pt}calculate\_distribution\textless-function(pars)\{\\
     \vspace{4pt}
     \# alpha and beta may be from a prior or posterior distribution\\
      alpha=pars[[1]]\\
      beta=pars[[2]]\\
      \vspace{4pt}
      \# Calculate the MLE of the mean\\
      mu\_numerator=alpha\\
      mu\_denominator=alpha+beta\\
      mu=mu\_numerator/mu\_denomenator\\
      \vspace{4pt}
      \# Calculate the MLE of the variance\\
      sig\_numerator=alpha\*beta\\
      den1=alpha+beta\\
      den2=den1$**$2\\
      sig\_denomenator=den1$*$den2\\
      sigma2=sig\_numerator/sig\_denomenator\\
      return(c(mu,sigma2))\}
    \end{em}
  \vspace{12pt}

  \vspace{12pt}
  \begin{em}
  \noindent \hspace{2pt}credible\_interval\textless-function(pars,p\_lower,p\_upper)\{\\
    \vspace{4pt}
    alpha=pars[[1]]\\
    beta=pars[[2]]\\
    lowCI=qbeta(p=p\_lower,shape1=alpha,shape2=beta)\\
    highCI=qbeta(p=p\_upper,shape1=alpha,shape2=beta)\\
    return(c(lowCI,highCI))\}
  \end{em}
  \vspace{12pt}

  To calculate the number of samples required to reach a given level of confidence that the probability of interaction between two species ($\lambda_{ij}$) that have not been observed co-occurring is below a threshold value:

  \vspace{12pt}
  \begin{em}
  \noindent \hspace{2pt}samples\_for\_threshold\textless-function(threshold,confidence,pars)\{\\
    \vspace{4pt}
    alpha=pars[[1]]\\
    beta=pars[[2]]\\
    n=seq(0,100,1)\\
    k=0\\ 
    cdf=pbeta(threshold,shape1=alpha,shape2=beta+n)\\
    samples=length(which(cdf\textless confidence))\\
    return(samples) \}
  \end{em}

\section*{Appendix S2: Details of \emph{Salix} data collection}

      The \emph{Salix}-galler-parasitoid meta-network dataset collected 
      by~\citet{Kopelke2017} consists of a single community type sampled across 
      Europe: willow (\emph{Salix}) species, willow-galling sawflies (
      Hymenoptera, Tenthredinidae, Nematinae, Euurina), and their natural 
      enemies (hymenopteran parasitoids and coleopteran, lepidopteran, dipteran, 
      and hymenopteran inquilines). The data were collected over 29 years 
      (1982-2010) at 374 unique locations across Europe ranging from Sicily to 
      the Arctic. Including repeated visits to a number of sites, the data set 
      contains 641 site-visits, each of which can be considered as a network in 
      its own right and an independent sample from which to build the 
      meta-network. The meta-network consists of 1,173 different interactions 
      between 52 \emph{Salix} nodes, 92 herbivore nodes, and 126 parasitoid 
      nodes. Interactions were determined by dissecting and rearing gall 
      inhabitants from 165,424 galls. Some sites were visited repeatedly, for a 
      total of 641 site-visits. We consider these to be repeated samples and 
      take the 374 unique sites as our sample size. The high spatiotemporal 
      resolution of this network and the unusually high sampling effort together 
      make this dataset particularly well suited for testing Bayesian approaches 
      to network structure.

\section*{Appendix S3: Obtaining prior interaction probabilities}

@@

      \citet{Barbour2016,Barbour2016Dryad} sampled 145 branches from 
      \emph{Salix hookeriana} of 26 genotypes. They recorded galls 
      made by four species of Cecidomyid midges on each branch. We 
      transformed these data to a binary genotype-galler network 
      matrix, where an entry $ij$ was 1 if galler $i$ was observed 
      on any branch of genotype $j$ and 0 otherwise. In total, this 
      network contained 75 realised galler-genotype interactions 
      out of a potential 104. One \emph{S. hookeriana} genotype did
      not interact with any gallers and was removed from the network.
      The R code used to extract this 
      network follows:

      \vspace{12pt}
      \begin{em}
      \noindent \hspace{2pt}prior\_web\_data=read.csv(``tree\_level\_interaxn\_all\_plants\_traits\_size.csv")\\
      \vspace{4pt}
      \noindent \hspace{2pt}\# Remove extra data on parasitoids, etc.\\
      \noindent \hspace{2pt}prior\_web\_data[,31:57]\textless-NULL\\
      \noindent \hspace{2pt}prior\_web\_data[,4:26]\textless-NULL\\
      \noindent \hspace{2pt}prior\_web\_data[,2]\textless-NULL\\
      \vspace{4pt}
      \noindent \hspace{2pt}\# Build a web with interactions=1 for any galler observed on any genotype\\
      \noindent \hspace{2pt}prior\_web=matrix(nrow=26,ncol=4)\\
      \noindent \hspace{2pt}for(r in 1:length(levels(prior\_web\_data\$Genotype)))\{\\
      \hspace{4pt}gen=levels(prior\_web\_data\$Genotype)[r]\\
      \hspace{4pt}subset=prior\_web\_data[which(prior\_web\_data\$Genotype==gen),]\\
      \hspace{4pt}for(c in 1:4)\{\\
      \hspace{8pt}if(sum(subset[,2+c])\textgreater0)\{\\
      \hspace{12pt}prior\_web[r,c]=1\\
      \hspace{8pt}\} else \{\\
      \hspace{12pt}prior\_web[r,c]=0\\
      \hspace{8pt}\}\}\}\\
      \vspace{4pt}
      \noindent \hspace{2pt}\# Remove non-interactive genotype to obtain the final web\\
      \noindent \hspace{2pt}prior\_web\textless-prior\_web[which(rowSums(prior\_web)\textgreater0),]
      \end{em}

@@

      \vspace{12pt}
      \clearpage

@@

      \vspace{12pt}
      \begin{em}
      \noindent \hspace{2pt}prior\_web\_data=read.csv(``tree\_level\_interaxn\_all\_plants\_traits\_size.csv")\\
      \noindent \hspace{2pt}prior\_web\_data \textless- prior\_web\_data[,8:26]\\
      \vspace{4pt}
      \noindent \hspace{2pt}\# Build a web with interactions=1 for any parasitoid on any galler\\
      \noindent \hspace{2pt}prior\_web=data.frame()\\
      \vspace{4pt}
      \noindent \hspace{2pt}\# each column contains all interactions for a \noindent \hspace{2pt}particular parasitoid-galler combination\\
      \noindent \hspace{2pt}for(column in 1:ncol(prior\_web\_data))\{\\
      \hspace{4pt}\# the column name gives the species which makes the gall,\\
      \hspace{4pt}\#followed by the species emerging from the gall\\
      \hspace{4pt}galler\_emerged=strsplit(colnames(prior\_web\_data)[column], "\_")\\
      \hspace{4pt}galler = galler\_emerged[[1]][1]\\
      \hspace{4pt}emerged = galler\_emerged[[1]][2]\\
      \vspace{4pt}
      \hspace{4pt}\# the strength of the interaction is the sum of the column\\
      \hspace{4pt}(number of galls across all branches sampled)\\
      \hspace{4pt}linkstrength = sum(prior\_web\_data[,column])\\
      \hspace{4pt}prior\_web[as.character(galler),as.character(emerged)] \textless- linkstrength\\
      \}\\
      \vspace{4pt}
      \noindent \hspace{2pt}\# Make the NAs into 0\\
      \noindent \hspace{2pt}prior\_web[is.na(prior\_web)] \textless- 0\\
      \vspace{4pt}
      \noindent \hspace{2pt}\# remove emerging gallers with no parasitoids\\
      \noindent \hspace{2pt}prior\_web\_para\_only \textless- prior\_web[,c(2:6,8:9,12)]\\
      \vspace{4pt}
      \noindent \hspace{2pt}\# convert interaction strengths to 1\\
      \noindent \hspace{2pt}prior\_web[which(prior\_web\textgreater0)] \textless- 1\\
      \noindent \hspace{2pt}write.csv(prior\_web\_para\_only, file=``prior\_web\_para\_only.csv", quote=F)\\
      \end{em}

@@

      \vspace{12pt}
      \begin{em}
      \noindent \hspace{2pt}\# Now get the degree distributions from the web\\
      \noindent \hspace{2pt}deg\_dist\_Salix=rowSums(prior\_web)/ncol(prior\_web)\\
      \vspace{4pt}
      \noindent \hspace{2pt}\# Remove one genotype that never interacted\\
      \noindent \hspace{2pt}deg\_dist\_Salix\textless-deg\_dist\_Salix[which(rowSums(prior\_web)\textgreater0)]\\
      \noindent \hspace{2pt}deg\_dist\_galler=colSums(prior\_web)/nrow(prior\_web)\\
      \vspace{4pt}
      \noindent \hspace{2pt}\# Interaction probabilities are the product of plant and galler probabilities\\
      \noindent \hspace{2pt}int\_probs=as.numeric(deg\_dist\_galler\%$*$\%t(deg\_dist\_Salix))\\
      \vspace{4pt}
      \noindent \hspace{2pt}\# And for the galler-parasitoid web:\\
      \noindent \hspace{2pt}deg\_dist\_parasitoid=rowSums(prior\_web\_para\_only)/ncol(prior\_web\_para\_only)\\
      \noindent \hspace{2pt}deg\_dist\_parasitized=colSums(prior\_web\_para\_only)/nrow(prior\_web\_para\_only)\\
      \noindent \hspace{2pt}gp\_int\_probs=as.numeric(deg\_dist\_parasitized\%$*$\%t(deg\_dist\_parasitoid))\\
      \end{em}

@@

    \bibliographystyle{ecollett} 
    \bibliography{MyCollection}
